---
title: "Skyine Analysis - AgZero Mix Experiment"
author: "Peter Levy"
output:
  html_document:
    number_sections: yes
    toc: no
  pdf_document:
    number_sections: yes
    toc: no
---

```{r render, echo = FALSE, eval = FALSE}
here::i_am("analysis/skyline_analysis_mix1.Rmd")
#renv::install("ggeffects")
library(here)
library(quarto)
library(rmarkdown)
quarto_render(here("analysis/skyline_analysis_mix1.Rmd"))
render(input = here("analysis", "skyline_analysis_mix1.Rmd"), output_file = here("analysis", "skyline_analysis_mix1.pdf"))
render(input = here("analysis", "skyline_analysis_mix1.Rmd"), output_file = here("analysis", "skyline_analysis_mix1.html"))
```

This document describes the statistical analysis of the skyline data from the AgZero mix (biochar and rockdust) experiment.

<!--- { startup -->
```{r startup, eval=TRUE, echo=FALSE, include=FALSE}
here::i_am("analysis/skyline_analysis_mix1.Rmd")
library(here)
library(targets)
library(ggplot2)
library(nlme)
library(mrds)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,
  results = "hold", fig.show = "hold")
Sys.setenv(TAR_PROJECT = "mix1")
plot_everything <- FALSE
```

## Distribution of the data
First we plot the distribution of the N2O fluxes to check for skewness.

```{r, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Histogram of N2O fluxes by treatment."}
dt_flux <- tar_read(dt_flux, store = here("_targets_mix1"))
dt <- dt_flux
flux_name <- "f_n2o"
mult <- 1000
dt[, f := get(flux_name) * mult]
dt[, trmt_id := as.factor(trmt_id)]
dt[, trmt_id := relevel(trmt_id, ref = "control")]
# table(dt[f < 0, trmt_id])
# 2.7 % are negative
# dim(dt[f < 0])[1] / dim(dt)[1] * 100
# fill negative fluxes with the minimum positive value
f_min <- dt[f > 0, min(f)]
# dt[f < 0, f := f_min]
dt <- dt[f > 0] # no, just subset


p <- ggplot(dt, aes(x = f)) +
  geom_histogram()
p <- p + facet_wrap(~ trmt_id)
p
```

The N2O flux data are very right-skewed. Usually a log-transformation converts this to something like a normal distribution.

```{r, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Histogram of log(N2O flux) by treatment."}
p <- ggplot(dt, aes(x = log(f))) +
  geom_histogram()
p <- p + facet_wrap(~ trmt_id)
p
```

Indeed, this is a lot more normal.
We need to include this consideration if we want to make summary statistics with uncertainty bounds.
We generally assume a normal distribution when we calculate confidence intervals and p values,
and here we can only justifiably do this with the log-transformed fluxes.

Out of interest, we can compare the pattern in the median and arithmetic mean across the treatments.
Differences come about when the data are skewed and asymmetrical.
(Fig below).

```{r, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Comparison of median and arithmetic mean flux across the treatments."}
gmean <- function(x) {
  exp(mean(log(x)))
}
dt_mean <- dt[, .(arithmetic_mean = mean(f), median = gmean(f)), by = trmt_id]
dt_mean <- melt(dt_mean, id.vars = "trmt_id", value = "mean_flux")
p <- ggplot(dt_mean, aes(x = trmt_id, y = mean_flux, fill = variable))
p <- p + geom_col(position = position_dodge(0.8), width = 0.7)
p <- p + facet_wrap(~ variable, scales = "free_y")
p
```
This suggests that the trends are not clear cut, as the pattern depends on which metric we use.
The relevant metric is the arithmetic mean (because the cumulative flux is the
arithmetic mean $\times$ time), but we cannot make direct inferences about that
because of the distribution of the data.


## Original analysis
We could ignore the skewness problem, and simply take the arimthmetic mean of the
flux in each chamber and analyse these in a standard linear model (~= ANOVA).

```{r, eval = FALSE, results = "hold", echo = TRUE}
dt_mean <- dt[, .(f = mean(f)), by = .(trmt_id, chamber_id)]
m_lm <- lm(f ~ trmt_id, data = dt_mean)
```

```{r, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Histogram of log(N2O flux) by treatment."}
dt_mean <- dt[, .(f = mean(f)), by = .(trmt_id, chamber_id)]

# Bar plots + jittered points + error bars
m_lm <- lm(f ~ trmt_id, data = dt_mean)
edf_lm <- ggpredict(m_lm, terms = c("trmt_id"))
df_lm <- as.data.table(edf_lm)
names(df_lm)[1] <- "trmt_id"
names(df_lm)[2] <- "f"

p <- ggplot(dt_mean, aes(trmt_id, f, colour = trmt_id))
p <- p + geom_hline(yintercept = 0)
p <- p +
  geom_col(
    data = df_lm,
    position = position_dodge(0.8),
    width = 0.7,
    fill = "white"
  )
p <- p + geom_jitter(position = position_jitter(0.2))
p <- p +
  geom_pointrange(
    data = df_lm,
    aes(ymin = conf.low, ymax = conf.high),
    colour = "black"
  )
p
```

This suggests "statistical significance" for the B2 treatment.

```{r, results = "hold", echo = TRUE}
summary(m_lm)
```
However, we have effectively just hidden the skewness problem: the intervals and
p values assume the chamber means are known quantities (with no uncertainty).
In fact, the chamber means are rather uncertain estimates when the data are very
skewed, but this has not been propagated into the statistical analysis.

## Analysis on the log-transformed data
To meet the standard assumptions, we need to do the analysis on the log-transformed data.
The complication this brings is that this estimates the median (or geometric mean)
rather than the arithmetic mean (which is what we want).
We can back-transform to get fluxes on the original scale, but an additional
calculation is needed to estimate the arithmetic mean.

To do this we fit a linear mixed model to the logged data, of the form below:

```{r, eval = FALSE, results = "hold", echo = TRUE}
m_lmer <- lmer(log(f) ~ trmt_id + (1 | chamber_id), data = dt)
```

Obtaining estimates of the treatment effects and intervals in his way, we then
apply the back-transformation and calculate the arithmetic means and intervals
with the standard equation for the lognormal distribution:

$$
\mu = exp(\mu_{\mathrm{log}} + \sigma_{\mathrm{log}}^2 / 2)
$$

where $\mu$ is the arithmetic mean on the original scale, $\mu_{\mathrm{log}}$
is the mean of the log-transformed data and
$\sigma_{\mathrm{log}}$ is the standard deviation of the log-transformed data.

```{r, eval = TRUE, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Effect of treatments on N2O flux (arithmetic means), showing 95% confidence intervals estimated from a linear mixed-effect model fitted to the log-transformed data. Results are back-transformed to the original scale and the arithmetic mean calculated according to the standard equation."}
dt_trmt <- dt[,
  .(amean = mean(f), geomean = gmean(f), f = exp(mean(log(f)) + sd(log(f))^2 / 2),
   bias_correction = sd(log(f))^2 / 2, n = .N),
  by = trmt_id
]
dt_trmt <- arrange(dt_trmt, factor(trmt_id, levels = levels(dt_trmt$trmt_id)))

dt_cham <- dt[,
  .(f = exp(mean(log(f)) + sd(log(f))^2 / 2), f_sd = sd(f, na.rm = TRUE), n = .N),
  by = .(trmt_id, chamber_id)
]
dt_cham[, chamber_id := as.factor(chamber_id)]

# Bar plots + jittered points + error bars
m_lmer <- lmer(log(f) ~ trmt_id + (1 | chamber_id), data = dt)
edf_lmer <- ggpredict(m_lmer, terms = c("trmt_id"))
df_lmer <- as.data.table(edf_lmer)
names(df_lmer)[1] <- "trmt_id"
names(df_lmer)[2] <- "f"
df_lmer[, bias_correction := dt_trmt$bias_correction]
df_lmer[, conf.int  := (conf.high - conf.low) / 2]
df_lmer[, f := exp(log(f) + bias_correction)]
df_lmer[, conf.low  := f - exp(log(conf.int) + bias_correction)]
df_lmer[, conf.high := f + exp(log(conf.int) + bias_correction)]

# plot(edf_lmer) + ylim(0, NA)
p <- ggplot(dt_cham, aes(trmt_id, f, colour = trmt_id))
p <- p + geom_hline(yintercept = 0)
p <- p +
  geom_col(
    data = df_lmer,
    position = position_dodge(0.8),
    width = 0.7,
    fill = "white"
  )
p <- p + geom_jitter(position = position_jitter(0.2))
p <- p +
  geom_errorbar(
    data = df_lmer,
    aes(ymin = conf.low, ymax = conf.high),
    width = 0.1,
    colour = "black"
  )
p
```

Using the more correct p values, the probablility that the treatment differences
are due to chance is relatively high (p > 0.4).
Although the patterns seen in the original analysis are the same, "statistical
significance" for the treatment differeces is not clear.

```{r, eval = TRUE, results = "hold", fig.show = "hold", echo = FALSE}
m_lme <- lme(log(f) ~ trmt_id, random = ~1 | chamber_id, data = dt)
summary(m_lme)
```
As model diagnostic checks, we should check that there is no trend in the residuals,
and that they are approximately normally distributed. Both of these seem resonable
in the plots below.

```{r, eval = TRUE, results = "hold", fig.show = "hold", echo = FALSE}
plot(m_lmer)
hist(resid(m_lmer))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "CO2 flux over time by treatment."}
summary(m_lmer)
dt[, f := get(flux_name) * mult]
m_lmer <- lmer(log(f) ~ trmt_id + (1 | chamber_id), data = dt)
predict_response(m_lmer)

# install.packages("emmeans")
library(emmeans)
emm.src <- emmeans(m_lmer, "trmt_id")
str(emm.src)
summary(emm.src, infer = TRUE, null = log(35))

summary(emm.src, infer = TRUE, type = "response")
plot(emm.src, type = "response")
plot(emm.src, type = "scale")

library(rstanarm)
m_stan <- stan_glmer(log(f) ~ trmt_id + (1 | chamber_id), data = dt)
plot(m_stan, "trace", par = 1:4)
pp_check(m_stan)

nd <- data.frame(trmt_id = levels(dt$trmt_id))
dt_post <- as.data.table(posterior_predict(m_stan, newdata = nd, re.form = NA))
names(dt_post) <- levels(dt$trmt_id)
dt_post <- melt(dt_post, variable = "trmt_id", value = "flux")
p <- ggplot(dt_post, aes(x = trmt_id, y = flux)) + geom_boxplot()
p

quantile(y_rep, c(0.025, 0.5, 0.975))
sapply(y_rep[1:4], quantile, c(0.025, 0.975))

library(nlme)
m_lme <- lme(log(f) ~ trmt_id, random = ~1| chamber_id, data = dt)
summary(m_lme)
anova(m_lme)
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "CO2 flux over time by treatment."}
tar_read(p_flux_co2, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, echo = FALSE, fig.cap = "CH4 flux over time by treatment."}
tar_read(p_flux_ch4, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, echo = FALSE, fig.cap = "N2O flux over time by treatment."}
tar_read(p_flux_n2o, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, echo = FALSE, fig.cap = "N2O flux versus temperature by treatment."}
tar_read(p_flux_n2o_T, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "N2O flux over time by treatment, showing N application dates as vertical black lines."}
tar_read(p_flux_n2o_with_Nappl, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Temperature data over time."}
tar_read(p_T, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "PPFD data over time."}
tar_read(p_Q, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "VWC data over time."}
tar_read(p_VWC, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Respiration response to temperature."}
tar_read(p_reco_T_response, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "GPP response to PPFD."}
tar_read(p_gpp_Q_response, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "GPP over time."}
tar_read(p_gpp_date, store = here("_targets_mix1"))
```

```{r, eval = plot_everything, results = "hold", fig.show = "hold", echo = FALSE, fig.cap = "Cummulative GPP, R and net flux over time."}
tar_read(p_cum, store = here("_targets_mix1"))
```
